Aspects of Dealing with Imperfect Data in Temporal Databases

### 2012-03-16 book/sources/fuzzyTemporalDatabases.tex
Still cool...

Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database. Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid time~\cite{garrido2009},~\cite{gal01}.

In the following subsection, a novel approach to representing uncertainty concerning valid time notions and a corresponding technique to query similar valid time indications in a valid-time relation are proposed. The presented proposal is based on concepts introduced in~\cite{Dubois88b} and on the framework proposed in~\cite{Pon11}.

\subsection{Handling Uncertainty in a Valid-time Relation}
A valid time indication usually takes the form of a time interval. Such a valid time interval can be described (and stored in a valid time relation record) using its boundaries (endpoints) or using one endpoint and the interval length. Usually, a valid-time interval is represented using its endpoints, which is also the approach adopted by the presented proposal.

@@

To represent valid time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}. To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first. In the presented proposal, the definition of possibilistic variables of~\cite{Pon11} is followed. In~\cite{Pon11}, a \emph{possibilistic variable} is defined as follows.

@@

Now, consider a set $R$, which contains single values (and not collections of values). When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set} \cite{Dubois88b}.

@@

In the presented proposal, a closed ill-known time interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted $\left[X, Y\right]$. Figure \ref{fig:interval} shows a closed ill-known time interval. %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research. Figure \ref{fig:interval} illustrates this.

@@

Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}. A comparison between the transformations in~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}. Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.

@@

In the presented proposal, ill-known valid time intervals will be used to represent valid time indications in a valid-time relation and to model the uncertainty these may contain.

To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}. Both concepts are treated in subsection \ref{iktc}.

Before ill-known time constraints can be introduced, another notion related to possibilistic variables shoud be paid attention to. In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false' \cite{Pon11}. Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ takes, is given by a possibility distribution $\pi_p$, then proposition $p$ can be seen as a possibilistic variable. In the presented proposal, the interest lies with the case where the proposition holds (denoted $p$ = $T$) and the possibility and necessity that $p$ = $T$ demand most attention. In the following sections, the following notations are used, based on previous notations:

@@

\subsubsection{\label{iktc}Ill-known Time Constraints}
The presented proposal contains a technique for evaluating user queries used to query a valid time relation in which the valid time indications are represented by ill-known time intervals. Part of this query evaluation technique relies on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}. These concepts are presented below. Following~\cite{Pon11}, an \emph{ill-known constraint} is defined as follows.

@@

The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are calculated using the following formulas~\cite{Pon11}.

@@

As mentioned before, the presented proposal deals with querying a valid time relation. In this subsection, the core of the presented proposal is described. First the particular structure of the relation is described, along with the nature and structure of its supposed contents. Next, the particular query structure is presented. Finally, the particular method for evaluating queries and for ranking the result records are presented.

@@

In the presented approach, for every record $r$, each part of a query $\tilde{Q} = \left( Q^{time}, Q \right)$ is evaluated independently:
\begin{itemize}
\item
The user preferences expressed in the non-temporal part $Q$, are evaluated, resulting in a satisfaction degree denoted $e_Q(r)$. The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
\item
The evaluation of the temporal demand expressed in the temporal part, $Q^{time} = \left( I,AR \right)$, depends on $AR$. A specific construct of ill-known constraints (cf. section~\ref{iktc}) is considered, depending on the Allen relation denoted by $AR$. The exact construct of constraints is an instantiation based on the formulas which can be found in table \ref{table:const}, for every possible value of $AR$. The form and capacity of these constraints are based on~\cite{Pon11}. Then, using equations \eqref{eq:pos} and \eqref{eq:nec}, the exact formulas to calculate the possibility $\Pos_{Q^{time}}(r)$ and the necessity $\Nec_{Q^{time}}(r)$ that record $r$ satisfies this construct of ill-known time constraints are derived from this construct of constraints. As mentioned, $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ denote the possibility, respectively the necessity that the considered record $r$ satisfies the construct of constraints corresponding to the temporal demand $Q^{time}$ and thus the possibility, respectively the necessity that $r$ satisfies $Q^{time}$.

@@

\subsection{Bipolarity in Temporal Databases}
\label{subsubsec:bipolarity}
Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need. This introduces the need for \emph{bipolar querying}, a technique to model both positive and negative user preferences in a database query. Sometimes positive and negative preferences are clearly symmetric, making it possible to derive one from the other. For example, a person may define the concept of `tall' as `1.80 meters or higher'. The negative would then be the opposite: not tall would be `anything less than 1.80 meters'. However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa. e.g., when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike. This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.

The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature \cite{DeTre2009}, \cite{Dubois2008}, \cite{Lacroix87}. In \cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences. However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences. However, the combination of bipolar querying and the use of imprecise query preferences in the context of temporal databases is not so well discussed in existing literature. A proposal for the bipolar querying of valid-time databases has been made by Billiet et al.~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. The model presented there deals with a fuzzy valid-time specification based on~\cite{garrido2009}.

@@

Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets \cite{Dubois02}.

>>>>

Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database. Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid-time~\cite{garrido2009},~\cite{gal01}.

In the following subsection, a novel approach to representing uncertainty concerning valid-time notions and a corresponding technique to query similar valid-time indications in a valid-time relation are proposed. The presented proposal is based on concepts introduced in~\cite{Dubois88b} and on the framework proposed in~\cite{Pon11}.

\subsection{Handling Uncertainty in a Valid-time Relation}
A valid-time indication usually takes the form of a time interval. Such a valid-time interval can be described (and stored in a valid-time relation record) using its boundaries (endpoints) or using one endpoint and the interval length. Usually, a valid-time interval is represented using its endpoints, which is also the approach adopted by the presented proposal.

@@

To represent valid-time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid-time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}. To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first. In the presented proposal, the definition of possibilistic variables of~\cite{Pon11} is followed. In~\cite{Pon11}, a \emph{possibilistic variable} is defined as follows.

@@

Now, consider a set $R$, which contains single values (and not collections of values). When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work~\cite{Dubois88b}.

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set}~\cite{Dubois88b}.

@@

In the presented proposal, a closed ill-known time interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted $\left[X, Y\right]$. Figure~\ref{fig:interval} shows a closed ill-known time interval. %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research. Figure \ref{fig:interval} illustrates this.

@@

Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}. A comparison between the transformations in~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}. Figure~\ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.

@@

In the presented proposal, ill-known valid-time intervals will be used to represent valid-time indications in a valid-time relation and to model the uncertainty these may contain.

To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid-time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}. Both concepts are treated in subsection~\ref{iktc}.

Before ill-known time constraints can be introduced, another notion related to possibilistic variables shoud be paid attention to. In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false'~\cite{Pon11}. Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ takes, is given by a possibility distribution $\pi_p$, then proposition $p$ can be seen as a possibilistic variable. In the presented proposal, the interest lies with the case where the proposition holds (denoted $p$ = $T$) and the possibility and necessity that $p$ = $T$ demand most attention. In the following sections, the following notations are used, based on previous notations:

@@

\subsubsection{\label{iktc}Ill-known Time Constraints}
The presented proposal contains a technique for evaluating user queries used to query a valid-time relation in which the valid-time indications are represented by ill-known time intervals. Part of this query evaluation technique relies on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}. These concepts are presented below. Following~\cite{Pon11}, an \emph{ill-known constraint} is defined as follows.

@@

The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not~\cite{Pon11}. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are calculated using the following formulas~\cite{Pon11}.

@@

As mentioned before, the presented proposal deals with querying a valid-time relation. In this subsection, the core of the presented proposal is described. First the particular structure of the relation is described, along with the nature and structure of its supposed contents. Next, the particular query structure is presented. Finally, the particular method for evaluating queries and for ranking the result records are presented.

@@

In the presented approach, for every record $r$, each part of a query $\tilde{Q} = \left( Q^{time}, Q \right)$ is evaluated independently:
\begin{itemize}
\item
The user preferences expressed in the non-temporal part $Q$, are evaluated, resulting in a satisfaction degree denoted $e_Q(r)$. The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
\item
The evaluation of the temporal demand expressed in the temporal part, $Q^{time} = \left( I,AR \right)$, depends on $AR$. A specific construct of ill-known constraints (cf. section~\ref{iktc}) is considered, depending on the Allen relation denoted by $AR$. The exact construct of constraints is an instantiation based on the formulas which can be found in table~\ref{table:const}, for every possible value of $AR$. The form and capacity of these constraints are based on~\cite{Pon11}. Then, using equations \eqref{eq:pos} and \eqref{eq:nec}, the exact formulas to calculate the possibility $\Pos_{Q^{time}}(r)$ and the necessity $\Nec_{Q^{time}}(r)$ that record $r$ satisfies this construct of ill-known time constraints are derived from this construct of constraints. As mentioned, $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ denote the possibility, respectively the necessity that the considered record $r$ satisfies the construct of constraints corresponding to the temporal demand $Q^{time}$ and thus the possibility, respectively the necessity that $r$ satisfies $Q^{time}$.

@@

\subsection{Bipolarity in Temporal Databases}
\label{subsubsec:bipolarity}
Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need. This introduces the need for \emph{bipolar querying}, a technique to model both positive and negative user preferences in a database query. Sometimes positive and negative preferences are clearly symmetric, making it possible to derive one from the other. For example, a person may define the concept of `tall' as `1.80 meters or higher'. The negative would then be the opposite: not tall would be `anything less than 1.80 meters'. However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa. E.g., when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike. This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.

The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature~\cite{DeTre2009},~\cite{Dubois2008},~\cite{Lacroix87}. In~\cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences. However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences. However, the combination of bipolar querying and the use of imprecise query preferences in the context of temporal databases is not so well discussed in existing literature. A proposal for the bipolar querying of valid-time databases has been made by Billiet et al.~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. The model presented there deals with a fuzzy valid-time specification based on~\cite{garrido2009}.

@@

Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets~\cite{Dubois02}.

### 2012-03-16 book/sources/fuzzyTimeDomain.tex
Still cool...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%INTRO
As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points \cite{Dyreson1994}. While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them. This phenomenon has been studied a.o. in the field of artificial intelligence~\cite{Tre97},~\cite{5151} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516},~\cite{nagypal2003},~\cite{Dev98}. An information system, however, cannot appeal to a similar interpretation functionality. Thus, many proposals have been concerned with the combination of time and imperfections in the context of information systems~\cite{nagypal2003}. In this section, some main concepts and issues concerning this combination are presented.

@@

\subsection{Imperfections in Temporal Relationships}
As the existence of temporal relationships allows to compare temporal notions, many approaches have been concerned with finding similar temporal relationships, able to support imperfections in the temporal information which is described by temporal notions or even by the temporal relationships themselves~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. These approaches are often based on Allen's operators~\cite{Allen83}.

%%%TODO

%Humans usually deals with time in a imprecise way. e.g. `A few days ago we received this package'. This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}. Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points. This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.

%When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:

%\begin{description}
%\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.

%\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way. e.g. `during summer', `late night'. Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} . e.g. `The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day. It is possible to set the following classification~\cite{Dev98}:

%\begin{itemize}
%\item
%Non-numerical indications. e.g. \emph{often, usually}.
%\item
%Approximative time indications. e.g. \emph{around 10 a.m.}
%\item
%Indications on temporal relations. e.g. \emph{after 12, during summer}.
%\end{itemize}

%\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
%\begin{itemize}
%\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
%\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer. e.g. A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m. or 6 a.m. The speaker, however knows exactly which one is meant. 
%\end{itemize}
%\end{description}

%Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%INTRO
As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points~\cite{Dyreson1994}. While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them. This phenomenon has been studied a.o. in the field of artificial intelligence~\cite{Tre97},~\cite{5151} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516},~\cite{nagypal2003},~\cite{Dev98}. An information system, however, cannot appeal to a similar interpretation functionality. Thus, many proposals have been concerned with the combination of time and imperfections in the context of information systems~\cite{nagypal2003}. In this section, some main concepts and issues concerning this combination are presented.

@@

In an information system, temporal information is usually related to facts or events~\cite{Chountas2000}. In light of this, a classification of temporal information can be considered, in which the following types of temporal information may be found:

\begin{itemize}
\item \emph{Definite temporal information}:
Definite temporal information contains information describing a situation in which all time indications associated with some fact are absolute time indications. The temporal information is precisely known.

\item \emph{Indefinite temporal information}~\cite{Dey1996}:
Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined. E.g., consider an event that in fact occured but it is not known exactly when.

\item \emph{Infinite temporal uncertain information}~\cite{Kabanza1990}:
Infinite temporal uncertain information contains information describing a situation in which an infinite number of time indications are associated with some fact. This is usually found in recurrent events like meetings. E.g., consider meetings that take place every Wednesday at noon. Some systems (usually with different information providers) may dispute the occurence and/or the duration of a fact.
\end{itemize}

\subsection{\label{subsec:representation}Representation of Imperfect Information}
As mentioned before, information systems may have to deal with time indications which contain vagueness. Even for some specific events or facts, the temporal indications may become imprecise. Therefore, a time point might be specified by means of a time interval of which the boundaries may not be precisely known. An example.

\begin{example}
Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saysing: \\
\begin{center}
\emph{`We will meet each other tomorrow around 10'}\\
\end{center}
The hearer will now usually instinctively agree that the appointment will be in e.g. the time interval between 9.55h and 10.05h.
\end{example}

The study of the semantics of `around' in temporal~\cite{Dev98} indications has shown that the size of the time interval associated with the imprecise specification of a time point depends on the distance with respect to the current time. E.g., consider now that the speaker is talking about something that happened \emph{`during last week'}, then the hearer would consider a time interval of more or less 10 days.

Some proposals~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003},~\cite{Chountas2000} conclude that the best representation for incomplete temporal knowledge is therefore based on time intervals, even if they refer to a fact that happen at a time point. This means that, as Allen proposed in~\cite{Allen83}, the primitive units (the chronons) in a time domain, used in an information system should be intervals.

In order to represent and manage uncertain temporal information properly, several theoretical frameworks have been proposed:

\begin{itemize}
\item \emph{Probability theory}:
Probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is usually employed when uncertainty concerning a time interval allows a probability to be associated to the time interval. The use of probability theory is very usual in logistics information systems. E.g., \emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.

\item \emph{Possibility theory}:
Using possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, a possibility degree is associated to the temporal fact or event. Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}. Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}. The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships.

\item \emph{Rough sets}:
Rough set theory~\cite{Pawlak1995} has been used to represent uncertainty in time intervals. The two dimensional representation of time intervals and the temporal relationships between them has been studied in~\cite{Qia09}.
\end{itemize}

\subsection{Imperfections in Temporal Relationships}
As the existence of temporal relationships allows to compare temporal notions, many approaches have been concerned with finding similar temporal relationships, able to support imperfections in the temporal information which is described by temporal notions or even by the temporal relationships themselves~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. These approaches are often based on Allen's operators~\cite{Allen83}. Example~\ref{ex:fuzzy-allen-relation} presents a short example concerning one of Allen's relationships.

\begin{example}
\label{ex:fuzzy-allen-relation}
Consider an event which takes place between time points $A$ and $B$. Thus, the event comprises time interval $[A,B]$ (this is visualized in part (1) of figure~\ref{fig:fuzzy-allen-relationship}). The classical Allen relationship `after' returns an interval $]B,\infty[$ as shown in part (2) of figure~\ref{fig:fuzzy-allen-relationship}. A fuzzified version of Allen's `after' operator is illustrated in part (3) of figure~\ref{fig:fuzzy-allen-relationship}. The comparison between two time intervals results in a possibility degree in the unit interval. The shape of the possibility distribution is shown in part (3) of figure~\ref{fig:fuzzy-allen-relationship}. Note that all the points strictly after the point $B$ results in a possibility degree of 1 whereas there is an area near the point $B$ in which the possibility degree runs smoothly between 0 and 1.

Consider now the interval given by $[C,D]$, illustrated in part (4) of figure~\ref{fig:fuzzy-allen-relationship}. The user wants to know if $[C,D]$ is after $[A,B]$. The crisp version of the `after' operator would return \emph{`no'} as an answer. The fuzzy version for the same operator would return \emph{`yes, with a possibility of 0.5'}.

\begin{figure}
\centering
\includegraphics[scale=0.5]{graphs/fuzzyAllen.eps}
\caption{Example for the Allen relationship \emph{`after'}. (1) The event bounded within time points $A$ and $B$. (2) The crisp version of the \emph{`after'} operator. (3) A fuzzy version of the after operator. (4) Another event, bounded within time points $[C,D]$.}
\label{fig:fuzzy-allen-relationship}
\end{figure}
\end{example}
